---
title: "RMET Analysis"
author: "Ellen Herschel, Colin McDaniel, Sherry Zhang"
date: "11/29/2021"
output:
  html_document:
    df_print: paged
    toc: yes
  pdf_document:
    toc: yes
---

## Load packages

```{r load-pkg, include=FALSE}
# Add additional packages you need
# install.packages("psych")
library(here)  # makes reading data more consistent
library(foreign)
library(tidyverse)  # for data manipulation and plotting
library(haven)  # for importing SPSS/SAS/Stata data
library(lme4)  # for multilevel analysis
library(glmmTMB)  # for longitudinal analysis
library(sjPlot)  # for plotting
library(modelsummary)  # for making tables
library(interactions)  # for interaction plots
library(nlme)
library(mice) # For missing data analysis
library(R.utils)
library(psych) # For plotting raw data
theme_set(theme_bw())  # Theme; just my personal preference
```

# Read in Data

```{r data-import}
# Import R data
alldata <- read.csv("~/Desktop/Coursework/Fall 2021/Multilevel Modeling/psyc575/Music_Sports_Empathy/Music_Sports_Empathy/Full Data Sets/sport_music_dat_mlm_long2.csv")

alldata <- read.csv("sport_music_dat_mlm_long2.csv")

widedata<- read.csv("~/Desktop/Coursework/Fall 2021/Multilevel Modeling/psyc575/Music_Sports_Empathy/Music_Sports_Empathy/Full Data Sets/sport_music_dat_mlm_wide.csv")

widedata<- read.csv("sport_music_dat_mlm_wide.csv")
```

#Factor, Get rid of NAs
```{r}
# Code Year and Group as factors
alldata$year <- factor(alldata$year)
alldata$group <- factor(alldata$group)

alldata$time<-as.integer(alldata$year)

# Remove NAs from data
alldata1<- drop_na(alldata, wasi_vci, group, rmet_percent_correct, wasi_age)
```

# Mean-center variables
```{r}
# Mean-center Age
alldata1 <- alldata1 %>% 
  group_by(id) %>%   # operate within individual
  mutate(age_pm = mean(wasi_age),   # create individual means (the same as `meanses`)
         age_pmc = wasi_age - age_pm) %>%   # cluster-mean centered
  ungroup()  # exit the "editing within groups" mode

alldata1 %>% 
  select(id, wasi_age, age_pm, age_pmc)

# Mean-center verbal ability
alldata1 <- alldata1 %>% 
  group_by(id) %>%   # operate within individual
  mutate(verbal_pm = mean(wasi_vci),   # create individual means (the same as `meanses`)
         verbal_pmc = wasi_vci - verbal_pm) %>%   # cluster-mean centered
  ungroup()  # exit the "editing within groups" mode

alldata1 %>% 
  select(id, wasi_vci, verbal_pm, verbal_pmc)


datasummary_skim(alldata1 %>%
                   select(rmet_percent_correct, time, wasi_age, group, wasi_vci ))
```


# Check for differences in baseline Age

```{r}

widedata$group <- factor(widedata$group)

agebase<-aov(wasi_age_Y1~group, widedata)
summary(agebase)

```

# Attrition Analysis
```{r}
# Replace incorrectly placed 0s with NA
alldata$rmet_percent_correct[alldata$rmet_percent_correct == 0] <- NA

# Examine missing data patterns
md.pattern(alldata[,c("rmet_percent_correct", "baseage", "wasi_vci")])

# Attrition analysis setup
alldata$miss_rmet_percent_correct <- ifelse(is.na(alldata$rmet_percent_correct), 1, 0)
alldata$miss_baseage <- ifelse(is.na(alldata$baseage), 1, 0)
alldata$miss_wasi_vci <- ifelse(is.na(alldata$wasi_vci), 1, 0)

# Test whether RMET is missing completely at random
t.test(alldata$baseage[alldata$miss_rmet_percent_correct==0], alldata$baseage[alldata$miss_rmet_percent_correct==1], alternative = "two.sided", conf.level = .95, var.equal = TRUE)
t.test(alldata$wasi_vci[alldata$miss_rmet_percent_correct==0], alldata$wasi_vci[alldata$miss_rmet_percent_correct==1], alternative = "two.sided", conf.level = .95, var.equal = TRUE)

# Test whether wasi_vci is missing completely at random
t.test(alldata$rmet_percent_correct[alldata$miss_wasi_vci==0], alldata$rmet_percent_correct[alldata$miss_wasi_vci==1], alternative = "two.sided", conf.level = .95, var.equal = TRUE)
t.test(alldata$baseage[alldata$miss_wasi_vci==0], alldata$baseage[alldata$miss_wasi_vci==1], alternative = "two.sided", conf.level = .95, var.equal = TRUE)
```

# Check assumptions

```{r assumptions}
alldata1 %>%
  # Select the variables to be used
  select(group, wasi_gender, rmet_percent_correct, baseage, wasi_age, wasi_vci) %>%
  psych::pairs.panels(ellipses = FALSE, cex = 0.2, cex.cor = 1)


```
The data for rmet_percent_correct, baseage, and wasi_vci (the continuous variables being included in the model) appear relatively normal.

# Write Down Model Equations 

Lv-1:
$$\mathrm{empathyRMET}_{ti} = \beta_{0i} + \beta_{1i}\mathrm{time}_{ti} + \beta_{2i}\mathrm{verbalpmc}_{ti} + e_{ti}$$
Lv-2:
$$\beta_{0i} = \gamma_{00} + \gamma_{01}\mathrm{baseage}_i + \gamma_{02}\mathrm{verbalpm_i} + \gamma_{03}\mathrm{group}_i+ u_{0i}$$
$$\beta_{1i} = \gamma_{10} + \gamma_{11}\mathrm{verbalpm}_i + \gamma_{12}\mathrm{group}_i +  u_{1i}$$
$$\beta_{2i} = \gamma_{20} + \gamma_{21}\mathrm{baseage}_i + \gamma_{22}\mathrm{group}_i + u_{2i}$$

# Baseline Model

```{r}
#model with just base interaction of year and group, and no random slopes
m0 <- glmmTMB(rmet_percent_correct ~ time + (1 | id),
    data = alldata1,
)

summary(m0)

```


# Add in and test covariates and random slopes

```{r}
#model 

m1 <- glmmTMB(rmet_percent_correct ~ time + group + (1 | id),
    data = alldata1,
)

summary(m1)

anova(m0, m1)
#m1 sig

#add interaction of year and group
m2 <- glmmTMB(rmet_percent_correct ~ time*group + (1 | id),
    data = alldata1,
)
summary(m2)

anova(m1, m2) #not sig, but this is what we want to test

#add more covariates
#verbal (person mean-centered)
m3 <- glmmTMB(rmet_percent_correct ~ time*group + verbal_pmc + (1 | id),
    data = alldata1,
)
summary(m3)

anova(m2, m3)
#is sig .055/2.

#verbal (person mean)
m4 <- glmmTMB(rmet_percent_correct ~ time*group + verbal_pmc+ verbal_pm + (1 | id),
    data = alldata1,
)
summary(m4)

anova(m3, m4)
#is sig (p val divided by 2)

# age
m5 <- glmmTMB(rmet_percent_correct ~ time*group + verbal_pmc+ verbal_pm + baseage + (1 | id),
    data = alldata1,
)
summary(m5)

anova(m4, m5)
# is sig

#gender
m6 <- glmmTMB(rmet_percent_correct ~ time*group + verbal_pmc+ verbal_pm + baseage + wasi_gender + (1 | id),
    data = alldata1,
)
summary(m6)

anova(m5, m6)
#not sig, gender doesnt seem to improve the model


#test for random slopes

#time
m5.1 <- glmmTMB(rmet_percent_correct ~ time*group + verbal_pmc+ verbal_pm + baseage +  (time | id),
    data = alldata1,
)
summary(m5.1)

anova(m5, m5.1)
#this is sig

#verbal
m5.2 <- glmmTMB(rmet_percent_correct ~ time*group + verbal_pmc+ verbal_pm + baseage + (time*verbal_pmc | id),
    data = alldata1,
        # The default optimizer did not converge; try optim
    control = glmmTMBControl(
        optimizer = optim,
        optArgs = list(method = "BFGS")
    )
)
summary(m5.2)
#model convergence problem?

anova(m5.1, m5.2)
#not sig

confint(m5.1, parm = "beta_")
```


# Here's our final model

```{r}
summary(m5.1)
```


# Coefficient Table (msummary)

```{r}
msummary(m5.1,
  stars=TRUE)

```


# Plot
```{r}
plot_model(m5.1,
    type = "pred", show.data = TRUE,
    title = "", dot.size = 0.5,
    jitter = 0.1
)

plot_model(m5.1,
    type = "pred", terms = "time",
    show.data = TRUE, jitter = 0.1,
    title = "Figure 1: Effect of time on empathy ",
    dot.size = 0.5
)

# Interaction plot (year:group)
interact_plot(m5.1,
              pred = "time",
              modx = "group",
              plot.points = TRUE,
              point.size = 0.5,
              point.alpha = 0.2,
              jitter = 0.02,
              x.label = "Year",
              y.label = "RMET Empathy")
```


Summary:
